---
title: "Generalized Linear Mixed Models for Moderator Analyses"
date: "2025-12-08"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    toc_depth: 4
    number_sections: true
    code_folding: hide
  word_document:
    toc: true
    toc_depth: 4
  pdf_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = FALSE, cache = FALSE)

# Set the root directory for the entire document (recommended way)
knitr::opts_knit$set(root.dir = dirname(rstudioapi::getActiveDocumentContext()$path))
```

# Preparation

```{r load-packages}
library(tidyverse)
library(lme4)
library(gt)
library(knitr)
library(brms)
library(rlang)
```

## Load Data

```{r load-data}
# load data
df_moderation <- read.csv2("../pre-processing/output/data_for_moderation_analyses.csv", dec = ".")

# Display the first rows to check the data
head(df_moderation)
```

## Prepare Trajectory Datasets

To analyse each symptom trajectory separately, we generate dedicated datasets in which the corresponding trajectory count and sample information are explicitly available. This step does not change the underlying data but clarifies the structure for later regression analyses and descriptive summaries. Each trajectory dataset also provides basic information about the total number of individuals and the number of contributing samples, grounding the modelling in a clear empirical context.

```{r trajectories}
# calculate Sample_Size_100
df_moderation <- df_moderation %>%
  mutate(Sample_Size_100 = Sample_Size / 100)

# correct labeling error
df_moderation$Health_First[df_moderation$MID == "M1356_2"] <- "No"

# add level "HP/first responder" to dataset
df_moderation <- df_moderation %>%
  mutate(Trauma_type = case_when(
    Health_First == "Yes" ~ "HP/first responder",
    TRUE ~ as.character(Trauma_type)
  ))

trajectory_setup <- function(df, traj) {
  
  # Column names
  perc_col <- paste0(traj, "_percentage")
  n_col    <- paste0(traj, "_n")
  
  df_out <- df %>%
    filter(!is.na(.data[[perc_col]])) %>%
    mutate(!!n_col := round((.data[[perc_col]] / 100) * Sample_Size))
  
  return(df_out)
}

df_Low <- trajectory_setup(df_moderation, "Low")
df_Decreasing <- trajectory_setup(df_moderation, "Decreasing")
df_Increasing <- trajectory_setup(df_moderation, "Increasing")
df_High <- trajectory_setup(df_moderation, "High")
df_Moderate <- trajectory_setup(df_moderation, "Moderate")

trajectory_descriptives <- function(df_traj, traj) {
  
  n_col <- paste0(traj, "_n")
  
  total_n <- sum(df_traj[[n_col]], na.rm = TRUE)
  unique_samples <- length(unique(df_traj$Study))
  
  additional_info <- data.frame(
    Description = c(
      paste0("Total number of individuals in ", traj, " trajectory"),
      "Number of unique samples"
    ),
    Value = format(c(total_n, unique_samples), big.mark = ",")
  )
  
  knitr::kable(
    additional_info,
    caption = paste("Additional Descriptives for", traj, "Trajectory Analysis")
  )
}
```

## Descriptive Statistics

We provide basic descriptive statistics of the included studies, without excluding any data due to missingness.

```{r descriptives}
# Calculate descriptives
total_number_of_samples <- nrow(df_moderation)
total_sample_size <- sum(df_moderation$Sample_Size, na.rm = TRUE)

# Create a summary table
summary_table <- data.frame(
  Description = c("Number of samples / cohorts", "Total sample size (sum of available Sample_Size)"),
  Value = format(c(total_number_of_samples, total_sample_size), big.mark = ",")
)

# Display as a table
knitr::kable(summary_table, caption = "Descriptive Statistics of the Dataset")
```
# Moderators

The analyses examine a broad set of potential moderators that may explain differences in the prevalence of specific PTSD symptom trajectories. Moderators are grouped into three conceptual domains, and are listed below using descriptive labels; categorical levels and reference categories appear in parentheses.

**Background variables**
Country income (Levels: High income countries; Mid-, low-income); Developmental age (Levels: Adult, Youth); Education (%); Having a partner (%); Mean age; Minorities (%); Percentage of women; US vs Non-US (Levels: Non-US, US)

**Trauma-related variables**
Civilian vs Military (Levels: Civilian, Military); Discrete vs non-discrete trauma (Levels: Non-discrete, Discrete); HP/First responder vs other (Levels: HP/First responder, Other); Occupational trauma (Levels: No, Yes); Trauma exposure (Levels: Interpersonal trauma, Non-interpersonal trauma); Trauma type (Levels: Combat, Injury, Natural disaster, HP/first responder, Other)

**Methodological variables**
Diagnostic Criteria (Levels: DSM-IV, DSM-5); Entropy; LGMM vs LCGA (Levels: LCGA, LGMM); Number of trajectories; PTSD symptom assessment (Levels: Interview, Self-reported scale); Quality score; Sample size / 100; Time between trauma and first assessment (months); Time between trauma and last assessment (months); Time point PTSD assessments (number); Time span of the study (months)

```{r substantive-groups-setup}
background_vars <- c(
  "Mean_age", "Developmental_age","Percentage_women", "Percentage_minority", "Percentage_partner", "High_education", "Location", "Location_US"
  )

trauma_vars <- c(
  "Discrete", "Trauma_type", "Military", "Occupational_trauma", "Trauma_exposure"
  )

method_vars <- c(
  "Sample_Size_100", "Grolts", "Entropy", "Scale_moderator", "Diagnostic_DSM", "Trajectory_analysis", "TP_assessments", "N_trajectories",
  "Trauma_TP1", "Trauma_TPX", "TP1_TPX"
  )

all_substantive_vars <- c(background_vars, trauma_vars, method_vars)

# Centralized Moderator Specifications
MOD_SPECS <- list(
  list(var="Percentage_women",      type="cont", label="Percentage of women"),
  list(var="Mean_age",              type="cont", label="Mean age"),
  list(var="Percentage_minority",   type="cont", label="Minorities (%)"),
  list(var="Percentage_partner",    type="cont", label="Having a partner (%)"),
  list(var="High_education",        type="cont", label="Education (%)"),
  list(var="TP_assessments",        type="cont", label="Time point PTSD assessments (number)"),
  list(var="N_trajectories",        type="cont", label="Number of trajectories"),
  list(var="Trauma_TP1",            type="cont", label="Time between trauma and first assessment (months)"),
  list(var="TP1_TPX",               type="cont", label="Time span of the study (months)"),
  list(var="Trauma_TPX",            type="cont", label="Time between trauma and last assessment (months)"),
  list(var="Entropy",               type="cont", label="Entropy"),
  list(var="Grolts",                type="cont", label="Quality score"),
  list(var="Sample_Size_100",       type="cont", label="Sample size / 100"),

  list(var="Diagnostic_DSM",        type="cat",  label="Diagnostic Criteria",
       keep=c("4","5"), ref="4",
       levels=c(`4`="DSM-IV", `5`="DSM-5")),

  list(var="Occupational_trauma",   type="cat", label="Occupational trauma",
       keep=c("No","Yes"), ref="No",
       levels=c(No="No", Yes="Yes")),

  list(var="Developmental_age",     type="cat", label="Developmental age",
       keep=c("Adult","Youth"), ref="Adult",
       levels=c(Adult="Adult", Youth="Youth")),

  list(var="Location",              type="cat", label="Country income",
       keep=c("High", "ML"), ref="ML",
       levels=c(High="High income countries",
                ML="Mid-, low- income")),

  list(var="Location_US",           type="cat", label="US vs Non-US", ref="Other",
       levels=c(Other="Non-US", US="US")),

  list(var="Trajectory_analysis",   type="cat", label="LGMM vs LCGA", ref="LCGA",
       levels=c(LCGA="LCGA", LGMM="LGMM")),

  list(var="Scale_moderator",        type="cat", label="PTSD symptom assessment",
       keep=c("Interview","Self"), ref="Interview",
       levels=c(Interview="Interview", Self="Self-reported scale")),

  list(var="Trauma_exposure",       type="cat", label="Trauma exposure",
       keep=c("Inter","Non"), ref="Inter",
       levels=c(Inter="Interpersonal trauma", Non="Non-interpersonal trauma")),

  llist(
    var="Trauma_type",
    type="cat",
    label="Trauma type",
    keep=c("Combat","Natural","Injury", "HP/first responder"), 
    ref="Injury",
    levels=c(
      Combat      = "Combat",
      Injury      = "Injury",
      Natural     = "Natural Disaster",
      "Natural "  = "Natural Disaster",   # map trailing-space version
      "HP/first responder" = "HP or First Responder"
    ),
    na_level = "Other"
  ),

  list(var="Military",              type="cat", label="Civilian vs Military", ref="No",
       levels=c(No="Civilian", Yes="Military")),

  list(var="Discrete",              type="cat", label="Discrete vs non-discrete trauma",
       keep=c("No","Yes"), ref="No",
       levels=c(No="Non-discrete", Yes="Discrete"))
)

MOD_LABELS <- purrr::map_chr(MOD_SPECS, "label") |>
  setNames(purrr::map_chr(MOD_SPECS, "var"))

CAT_SPECS <- MOD_SPECS[sapply(MOD_SPECS, `[[`, "type") == "cat"]
names(CAT_SPECS) <- sapply(CAT_SPECS, `[[`, "var")

LEVEL_MAP <- purrr::map_dfr(MOD_SPECS, \(spec) {
  
  # Initialize the level data frame
  df_levels <- tibble::tibble(
    moderator = character(),
    orig      = character(),
    pretty    = character(),
    stan      = character()
  )
  
  # 1. Handle explicit levels
  if (!is.null(spec$levels)) {
    df_levels <- tibble::tibble(
      moderator = spec$var,
      orig      = names(spec$levels),
      pretty    = unname(spec$levels)
    )
  }
  
  # 2. Handle na_level (missing values)
  if (!is.null(spec$na_level)) {
    # map NA (the original missing value) to the pretty na_level.
    df_na <- tibble::tibble(
      moderator = spec$var,
      orig      = NA_character_, 
      pretty    = spec$na_level
    )
    df_levels <- dplyr::bind_rows(df_levels, df_na)
  }
  
  if (nrow(df_levels) == 0) return(NULL)
  
  # Add the 'stan' cleaned version
  df_levels <- df_levels %>%
    dplyr::mutate(stan = gsub("[^A-Za-z0-9]", "", pretty))
  
  return(df_levels)
})

```

Below we present a summary of all moderators included in the analysis, across the full dataset, as well as trajectory specific. The table presents overall descriptive statistics for each moderator variable and compares how these moderators are distributed across the five identified trajectory classes (Low, Decreasing, Increasing, High, and Moderate). Continuous moderators are summarized using the mean, standard deviation, and observed range. Categorical moderators show the proportion of studies falling into each category overall and within each trajectory class. Moderators are grouped into background characteristics, trauma characteristics, and methodological characteristics to facilitate interpretation.

```{r moderator-summary-gt, message=FALSE}
make_combined_mod_table <- function(df_mod,
                                    MOD_SPECS,
                                    LEVEL_MAP,
                                    background_vars = NULL,
                                    trauma_vars     = NULL,
                                    method_vars     = NULL) {

  trajectory_cols <- c("Low","Decreasing","Increasing","High","Moderate")

  # 0. Make sure trajectory columns are 0/1 numerics
  for (traj in trajectory_cols) {
    if (traj %in% names(df_mod)) {
      df_mod[[traj]] <- as.numeric(df_mod[[traj]] %in% c(1, "1", TRUE, "TRUE"))
      df_mod[[traj]][is.na(df_mod[[traj]])] <- 0
    }
  }

  # 1. Clean LEVEL_MAP
  LEVEL_MAP <- LEVEL_MAP %>%
    mutate(orig = trimws(orig)) %>%
    distinct(moderator, pretty, .keep_all = TRUE)

  # 2. assign group (Background / Trauma / Method / Other)
  get_group <- function(var) {
    if (!is.null(background_vars) && var %in% background_vars) {
      "Background characteristics"
    } else if (!is.null(trauma_vars) && var %in% trauma_vars) {
      "Trauma characteristics"
    } else if (!is.null(method_vars) && var %in% method_vars) {
      "Methodological characteristics"
    } else {
      "Other"
    }
  }

  # 3. recode a categorical moderator to pretty labels
  recode_categorical <- function(x_raw, spec, spec_map) {

    x <- x_raw

    # restrict to 'keep' values if specified
    if (!is.null(spec$keep)) {
      x[!(x %in% spec$keep)] <- NA
    }

    x <- trimws(as.character(x))

    non_na_map <- spec_map %>% filter(!is.na(orig))
    na_map     <- spec_map %>% filter(is.na(orig))

    if (nrow(non_na_map) > 0) {
      recode_map <- setNames(non_na_map$pretty, non_na_map$orig)
      x <- dplyr::recode(x, !!!recode_map, .default = NA_character_)
    } else {
      x <- as.character(x)
    }

    if (nrow(na_map) >= 1) {
      x[is.na(x)] <- na_map$pretty[1]
    }

    x
  }

  # 4. OVERALL SUMMARIES
  overall_rows <- map_dfr(MOD_SPECS, function(spec) {

    var   <- spec$var
    label <- spec$label
    type  <- spec$type
    x_raw <- df_mod[[var]]
    group <- get_group(var)

    if (type == "cont") {

      Mean <- mean(x_raw, na.rm = TRUE)
      SD   <- sd(x_raw, na.rm = TRUE)
      Min  <- min(x_raw, na.rm = TRUE)
      Max  <- max(x_raw, na.rm = TRUE)

      tibble(
        Group          = group,
        Moderator      = label,
        Type           = "Continuous",
        Level          = "",
        OverallSummary = glue("{round(Mean, 2)} (SD = {round(SD, 2)}), {round(Min, 2)}–{round(Max, 2)}")
      )

    } else {

      spec_map <- LEVEL_MAP %>% filter(moderator == var)

      # recode to pretty labels
      x_rec <- recode_categorical(x_raw, spec, spec_map)

      # full list of level labels (pretty)
      all_levels <- tibble(Level = spec_map$pretty)

      tab_df <- tibble(Level = x_rec) %>%
        filter(!is.na(Level)) %>%
        count(Level, name = "N") %>%
        mutate(P = N / sum(N))

      full_df <- all_levels %>%
        left_join(tab_df, by = "Level") %>%
        mutate(
          N = replace_na(N, 0L),
          P = replace_na(P, 0),
          OverallSummary = sprintf("%d (%.1f%%)", N, P * 100)
        )

      tibble(
        Group          = group,
        Moderator      = label,
        Type           = "Categorical",
        Level          = full_df$Level,
        OverallSummary = full_df$OverallSummary
      )
    }
  })

  # 5. TRAJECTORY SUMMARIES
  traj_rows <- map_dfr(MOD_SPECS, function(spec) {

    var   <- spec$var
    label <- spec$label
    type  <- spec$type

    if (type == "cont") {

      mean_vals <- sapply(trajectory_cols, function(traj) {
        vals <- df_mod[[var]][df_mod[[traj]] == 1]
        mean(vals, na.rm = TRUE)
      })

      sd_vals <- sapply(trajectory_cols, function(traj) {
        vals <- df_mod[[var]][df_mod[[traj]] == 1]
        sd(vals, na.rm = TRUE)
      })

      formatted <- mapply(
        function(m, s) {
          if (is.na(m)) return("–")
          glue("{round(m, 1)} (SD = {round(s, 1)})")
        },
        mean_vals, sd_vals
      )

      tibble(
        Moderator = label,
        Level     = "",
        !!!setNames(as.list(formatted), trajectory_cols)
      )

    } else {

      spec_map <- LEVEL_MAP %>% filter(moderator == var)
      x_rec    <- recode_categorical(df_mod[[var]], spec, spec_map)

      map_dfr(spec_map$pretty, function(lev_pretty) {

        level_idx <- x_rec == lev_pretty

        row_vals <- sapply(trajectory_cols, function(traj) {

          traj_idx   <- df_mod[[traj]] == 1
          numerator  <- sum(level_idx & traj_idx, na.rm = TRUE)
          denominator <- sum(traj_idx, na.rm = TRUE)

          if (denominator == 0) {
            return("–")
          } else {
            perc <- 100 * numerator / denominator
            sprintf("%.1f%%", perc)
          }
        })

        tibble(
          Moderator = label,
          Level     = lev_pretty,
          !!!as.list(row_vals)
        )
      })
    }
  })

  # 6. Merge overall + trajectory summaries
  final_df <- overall_rows %>%
    left_join(traj_rows, by = c("Moderator", "Level"))

  # order groups and moderators according to MOD_SPECS
  group_levels <- c(
    "Background characteristics",
    "Trauma characteristics",
    "Methodological characteristics",
    "Other"
  )

  mod_order <- map_chr(MOD_SPECS, "label")

  final_df <- final_df %>%
    mutate(
      Group = factor(Group, levels = group_levels),
      Moderator = factor(Moderator, levels = mod_order)
    ) %>%
    arrange(Group, Moderator)

  # 7. Render GT table
  gt(final_df, groupname_col = "Group") %>%
    cols_label(
      Moderator      = md("**Moderator**"),
      Type           = md("**Type**"),
      Level          = md("**Level**"),
      OverallSummary = md("**Overall Summary**"),
      Low            = md("**Low**"),
      Decreasing     = md("**Decreasing**"),
      Increasing     = md("**Increasing**"),
      High           = md("**High**"),
      Moderate       = md("**Moderate**")
    ) %>%
    cols_align(
      align = "left",
      columns = c(Moderator, Type, Level, OverallSummary)
    ) %>%
    cols_align(
      align = "right",
      columns = trajectory_cols
    ) %>%
    tab_spanner(
      label   = "Trajectory distributions (% within trajectory)",
      columns = trajectory_cols
    ) %>%
    tab_header(
      title = md("**Moderator Summary with Trajectory Distributions**"),
      subtitle = "Categorical moderators show % within each trajectory; continuous moderators show mean (SD)"
    ) %>%
    opt_row_striping() %>%
    cols_hide(Group) %>%
    tab_options(
      table.font.size              = 14,
      table.font.names             = "Times New Roman",
      column_labels.background.color = "#f0f0f0",
      column_labels.font.weight    = "bold",
      data_row.padding             = px(4),
      table.width                  = pct(100)
    ) %>%
    opt_table_lines() %>%
    # Bold the first occurrence of each moderator
    tab_style(
      style = cell_text(weight = "bold"),
      locations = cells_body(
        columns = Moderator,
        rows = !duplicated(Moderator)
      )
    ) %>%
    # Blank repeated moderator names for cleaner look
    text_transform(
      locations = cells_body(columns = Moderator),
      fn = function(x) {
        x[duplicated(x)] <- ""
        x
      }
    )
}

summary_tbl <- make_combined_mod_table(
  df_moderation,
  MOD_SPECS,
  LEVEL_MAP,
  background_vars = background_vars,
  trauma_vars     = trauma_vars,
  method_vars     = method_vars
)
summary_tbl
```

# Model 1: Univariate GLMMs 

The first modelling stage investigates each moderator separately using a generalized linear mixed model with a binomial likelihood and a random intercept for study. 
This univariate approach provides an initial screening of simple associations between each moderator and the prevalence of a given trajectory. 

For continuous moderators, the model describes how the expected trajectory prevalence changes with a one-unit increase in the predictor. For categorical moderators, it compares the prevalence between the reference category and each alternative category. These univariate models offer clear descriptive insights before moving to the more complex multivariate analyses.

For each trajectory (Low, Decreasing, Increasing, High, and Moderate) we repeat the univariate analyses, accompanied by descriptive summaries and correlation matrices of the continuous predictors for that trajectory. This allows us to examine how predictors relate to each specific symptom course and whether certain patterns are consistent across trajectories.

```{r mods-setup-fitting}
# Consistent Optimizer
ctrl <- lme4::glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))

# add Success/Failure columns
add_binom_cols <- function(data, success_col, denom_col) {
  data %>%
    dplyr::mutate(
      Success = .data[[success_col]],
      Failure = .data[[denom_col]] - Success
    )
}

# extract a p-value from a model summary
get_pval <- function(coefs, var) {
  p_cols <- intersect(colnames(coefs), c("Pr(>|z|)", "Pr(>|t|)", "Pr(>|Chisq|)"))
  if (length(p_cols) && var %in% rownames(coefs)) {
    return(unname(coefs[var, p_cols[1]]))
  }
  return(NA_real_)
}

# Fit continuous model
fit_cont <- function(data, var, success_col, denom_col) {
  d <- data %>%
    dplyr::filter(!is.na(.data[[var]])) %>%
    add_binom_cols(success_col, denom_col)

  m <- lme4::glmer(
    stats::as.formula(paste0("cbind(Success, Failure) ~ ", var, " + (1|Study)")),
    data = d,
    family = binomial,
    control = ctrl   # uses global ctrl
  )

  fe <- lme4::fixef(m)
  b0 <- fe["(Intercept)"]
  b1 <- fe[var]
  p0 <- plogis(b0)
  s  <- summary(m)$coefficients

  tibble::tibble(
    predictor = var,
    Samples   = stats::nobs(m),
    prevalence_at_0_pct = 100 * p0,
    slope = b1,
    marginal_delta_pp_per_unit = b1 * p0 * (1 - p0) * 100,
    p_value = get_pval(s, var),
    type="cont"
  )
}


# Fit categorical model
fit_cat <- function(data, var, success_col, denom_col) {
  spec <- CAT_SPECS[[var]]
  if (is.null(spec)) return(tibble::tibble())

  d <- data %>%
    dplyr::filter(!is.na(.data[[var]])) %>%
    { if (!is.null(spec$keep)) dplyr::filter(., .data[[var]] %in% spec$keep) else . } %>%
    add_binom_cols(success_col, denom_col) %>%
    dplyr::mutate(
      !!var := forcats::fct_drop(forcats::as_factor(.data[[var]])),
      !!var := if (!is.null(spec$ref) && spec$ref %in% levels(.data[[var]]))
                 forcats::fct_relevel(.data[[var]], spec$ref)
               else .data[[var]]
    )

  if (nlevels(d[[var]]) < 2) return(tibble::tibble())

  m <- lme4::glmer(
    stats::as.formula(paste0("cbind(Success, Failure) ~ ", var, " + (1|Study)")),
    data = d,
    family = binomial,
    control = ctrl   # uses global ctrl
  )

  fe <- lme4::fixef(m)
  s  <- summary(m)$coefficients
  lvls <- levels(d[[var]])

  add <- purrr::map_dbl(lvls, \(L) {
    nm <- paste0(var, L)
    if (nm %in% names(fe)) fe[[nm]] else 0
  })

  p <- plogis(fe["(Intercept)"] + add)
  p_ref <- p[1]

  tibble::tibble(
    moderator      = var,
    n              = stats::nobs(m),
    ref_level      = lvls[1],
    level          = lvls,
    prevalence_pct = 100 * p,
    diff_vs_ref_pp = 100 * (p - p_ref),
    p_value        = purrr::map_dbl(lvls, \(L) get_pval(s, paste0(var, L))),
    type="cat"
  )
}

# Run a set of variables on a trajectory dataset
run_group <- function(df_traj, vars, success_col, denom_col = "Sample_Size") {

  results <- purrr::map_dfr(vars, \(v) {
    spec <- purrr::keep(MOD_SPECS, ~ .x$var == v)[[1]]

    if (spec$type == "cont") {
      fit_cont(df_traj, v, success_col, denom_col)
    } else {
      fit_cat(df_traj, v, success_col, denom_col)
    }
  })

  list(
    cont = dplyr::filter(results, type=="cont"),
    cat  = dplyr::filter(results, type=="cat")
  )
}
```

A summary table for each group of predictors shows how the univariate models perform per conceptual domain (i.e., Background, Trauma-related, and Methodological variables).
We also provide a summary table for all univariate models of the whole trajectory, also grouped into conceptual domains.

```{r mods-setup-table}
# P-value formatting
fmt_p <- function(p) ifelse(is.na(p), "", ifelse(p < .001, "< .001", sprintf("%.3f", p)))

sig <- function(p) dplyr::case_when(is.na(p) ~ "", p < .001 ~ "***", p < .01 ~ "**", p < .05 ~ "*", TRUE ~ "")

run_mods <- function(df, vars, success_col, denom_col="Sample_Size") {
  purrr::map_dfr(vars, \(v) {
    spec <- purrr::keep(MOD_SPECS, ~ .x$var == v)[[1]]
    if (spec$type == "cont") fit_cont(df, v, success_col, denom_col)
    else fit_cat(df, v, success_col, denom_col)
  })
}

# Table Rendering Function
.na_label <- "-"

make_group_table <- function(group_res, title, na_label = .na_label) {

  # Continuous
  cont_tbl <- if (!is.null(group_res$cont) && nrow(group_res$cont) > 0) {

    group_res$cont %>%
      dplyr::mutate(
        Predictor            = dplyr::recode(predictor, !!!MOD_LABELS, .default = predictor),
        k_shared             = as.integer(Samples),
        Ref_Prevalence_Cont  = sprintf("%.2f%%", prevalence_at_0_pct),
        Delta_Marginal_Cont  = paste0(round(marginal_delta_pp_per_unit, 2)),
        p_value_shared       = paste0(fmt_p(p_value),
                                      ifelse(is.na(p_value), "", paste0(" ", sig(p_value)))),
        Ref_Name_Cat         = na_label,
        Ref_Prevalence_Cat   = na_label,
        Comp_Name_Cat        = na_label,
        Comp_Prevalence_Cat  = na_label,
        Delta_Marginal_Cat   = na_label
      ) %>%
      dplyr::select(dplyr::all_of(UNIFIED_COLS))

  } else {
    .empty_unified_tbl()
  }

  # Categorical
  cat_tbl <- if (!is.null(group_res$cat) && nrow(group_res$cat) > 0) {

    ref_prev <- group_res$cat %>%
      dplyr::filter(level == ref_level) %>%
      dplyr::select(moderator, ref_level, ref_prev = prevalence_pct)

    group_res$cat %>%
      dplyr::filter(level != ref_level) %>%
      dplyr::left_join(ref_prev, by = c("moderator","ref_level")) %>%
      dplyr::left_join(LEVEL_MAP, by = c("moderator","ref_level" = "orig")) %>%
      dplyr::rename(pretty_ref = pretty) %>%
      dplyr::left_join(LEVEL_MAP, by = c("moderator","level" = "orig")) %>%
      dplyr::rename(pretty_cmp = pretty) %>%
      dplyr::mutate(
        Predictor            = dplyr::recode(moderator, !!!MOD_LABELS, .default = moderator),
        k_shared             = as.integer(n),
        Ref_Prevalence_Cont  = na_label,
        Delta_Marginal_Cont  = na_label,
        Ref_Name_Cat         = dplyr::coalesce(pretty_ref, ref_level),
        Ref_Prevalence_Cat   = sprintf("%.1f%%", ref_prev),
        Comp_Name_Cat        = dplyr::coalesce(pretty_cmp, level),
        Comp_Prevalence_Cat  = sprintf("%.1f%%", prevalence_pct),
        Delta_Marginal_Cat   = sprintf("%.2f", diff_vs_ref_pp),
        p_value_shared       = paste0(fmt_p(p_value),
                                      ifelse(is.na(p_value), "", paste0(" ", sig(p_value))))
      ) %>%
      dplyr::select(dplyr::all_of(UNIFIED_COLS))

  } else {
    .empty_unified_tbl()
  }

  # Bind + render
  out <- dplyr::bind_rows(cont_tbl, cat_tbl)

  gt::gt(out) %>%
    gt::tab_header(title = enc2utf8(title)) %>%
    gt::cols_label(
      Predictor = gt::md("**Predictor**"),
      k_shared = gt::md("k"),
      p_value_shared = gt::md("p-value"),
      Ref_Prevalence_Cont = gt::md("Relative<br>prevalence at 0%"),
      Delta_Marginal_Cont = gt::md("Marginal Δ"), 
      Ref_Name_Cat = gt::md("Reference (Name)"),
      Ref_Prevalence_Cat = gt::md("Reference Prevalence"),
      Comp_Name_Cat = gt::md("Comparison (Name)"),
      Comp_Prevalence_Cat = gt::md("Comparison Prevalence"),
      Delta_Marginal_Cat = gt::md("Marginal Δ")
    ) %>%
    gt::tab_spanner(
      label = gt::md("**Continuous Moderators**"),
      columns = c(Ref_Prevalence_Cont, Delta_Marginal_Cont),
      id = "cont_group"
    ) %>%
    gt::tab_spanner(
      label = gt::md("**Categorical Moderators**"),
      columns = c(Ref_Name_Cat, Ref_Prevalence_Cat, Comp_Name_Cat,
                  Comp_Prevalence_Cat, Delta_Marginal_Cat),
      id = "cat_group"
    ) %>%
    gt::cols_align("left",  columns = Predictor) %>%
    gt::cols_align("center", columns = -Predictor) %>%
    gt::cols_width(Predictor ~ gt::px(180),
                   k_shared ~ gt::px(50),
                   p_value_shared ~ gt::px(70)) %>%
    gt::tab_style(
      style = gt::cell_text(weight = "bold"),
      locations = gt::cells_column_labels(columns = gt::everything())
    ) %>%
    gt::tab_options(
      table.width = gt::pct(100),
      data_row.padding = gt::px(4),
      column_labels.padding = gt::px(6),
      table.font.size = gt::px(13),
      table.font.names = "Times New Roman"
    )
}

UNIFIED_COLS <- c(
  "Predictor", "k_shared",
  "Ref_Prevalence_Cont", "Delta_Marginal_Cont", "p_value_shared",
  "Ref_Name_Cat", "Ref_Prevalence_Cat", "Comp_Name_Cat", "Comp_Prevalence_Cat",
  "Delta_Marginal_Cat"
)

.empty_unified_tbl <- function() {
  tibble::tibble(
    Predictor            = character(),
    k_shared             = integer(),
    Ref_Prevalence_Cont  = character(),
    Delta_Marginal_Cont  = character(),
    p_value_shared       = character(),
    Ref_Name_Cat         = character(),
    Ref_Prevalence_Cat   = character(),
    Comp_Name_Cat        = character(),
    Comp_Prevalence_Cat  = character(),
    Delta_Marginal_Cat   = character()
  )
}
```

## 🚀 Low Symptom Trajectory

### Descriptives

We provide additional descriptives based on the Low trajectory variable, using the full dataset.

```{r Low-descriptives}
trajectory_descriptives(df_Low, "Low")
```

Below, we fit univariate models of predictors for the Low Trajectory grouped by conceptual domains: Background, Trauma-related, and Methodological variables.
For continuous moderators, the model describes how the expected trajectory prevalence changes with a one-unit increase in the predictor. 
For categorical moderators, it compares the prevalence between the reference category and each alternative category.
Differences (Marginal Δ) are significant at α = .05. 

```{r Low-mods-fitting}
# Run the three groups for this trajectory
bg_res_Low <- run_group(df_Low, success_col = "Low_n", denom_col = "Sample_Size", vars = background_vars)
tr_res_Low <- run_group(df_Low, success_col = "Low_n", denom_col = "Sample_Size", vars = trauma_vars)
md_res_Low <- run_group(df_Low, success_col = "Low_n", denom_col = "Sample_Size", vars = method_vars)
all_Low_res <- run_group(df_Low, success_col = "Low_n", denom_col = "Sample_Size", vars = all_substantive_vars)
```

```{r Low-mods-tables}
make_group_table(bg_res_Low, "Low Symptom Trajectory: Socio-Demographic Moderators")
make_group_table(tr_res_Low, "Low Symptom Trajectory: Trauma-Related Moderators")
make_group_table(md_res_Low, "Low Symptom Trajectory: Methodological Moderators")
```

Below is a summary of univariate models of all predictor variables for the Low Trajectory.

```{r Low-mods-tables-summary}
make_group_table(all_Low_res, "Low Symptom Trajectory: All Moderators")
```

## 🌿 Decreasing Symptom Trajectory

### Descriptives

We provide additional descriptives based on the Decreasing trajectory variable, using the full dataset.

```{r Decreasing-descriptives}
trajectory_descriptives(df_Decreasing, "Decreasing")
```

Below, we fit univariate models of predictors for the Decreasing Trajectory grouped by conceptual domains: Background, Trauma-related, and Methodological variables.
For continuous moderators, the model describes how the expected trajectory prevalence changes with a one-unit increase in the predictor. 
For categorical moderators, it compares the prevalence between the reference category and each alternative category.
Differences (Marginal Δ) are significant at α = .05. 

```{r Decreasing-mods-fitting}
# Run the three groups for this trajectory
bg_res_Decreasing <- run_group(df_Decreasing, success_col = "Decreasing_n", denom_col = "Sample_Size", vars = background_vars)
tr_res_Decreasing <- run_group(df_Decreasing, success_col = "Decreasing_n", denom_col = "Sample_Size", vars = trauma_vars)
md_res_Decreasing <- run_group(df_Decreasing, success_col = "Decreasing_n", denom_col = "Sample_Size", vars = method_vars)
all_Decreasing_res <- run_group(df_Decreasing, success_col = "Decreasing_n", denom_col = "Sample_Size", vars = all_substantive_vars)
```

```{r Decreasing-mods-tables}
make_group_table(bg_res_Decreasing, "Decreasing Symptom Trajectory: Socio-Demographic Moderators")
make_group_table(tr_res_Decreasing, "Decreasing Symptom Trajectory: Trauma-Related Moderators")
make_group_table(md_res_Decreasing, "Decreasing Symptom Trajectory: Methodological Moderators")
```

Below is a summary of univariate models of all predictor variables for the Decreasing Trajectory.

```{r Decreasing-mods-tables-summary}
make_group_table(all_Decreasing_res, "Decreasing Symptom Trajectory: All Moderators")
```

## ⚠️ Increasing Symptom Trajectory

### Descriptives

We provide additional descriptives based on the Increasing trajectory variable, using the full dataset.

```{r Increasing-descriptives}
trajectory_descriptives(df_Increasing, "Increasing")
```

Below, we fit univariate models of predictors for the Increasing Trajectory grouped by conceptual domains: Background, Trauma-related, and Methodological variables.
For continuous moderators, the model describes how the expected trajectory prevalence changes with a one-unit increase in the predictor. 
For categorical moderators, it compares the prevalence between the reference category and each alternative category.
Differences (Marginal Δ) are significant at α = .05. 

```{r Increasing-mods-fitting}
# Run the three groups for this trajectory
bg_res_Increasing <- run_group(df_Increasing, success_col = "Increasing_n", denom_col = "Sample_Size", vars = background_vars)
tr_res_Increasing <- run_group(df_Increasing, success_col = "Increasing_n", denom_col = "Sample_Size", vars = trauma_vars)
md_res_Increasing <- run_group(df_Increasing, success_col = "Increasing_n", denom_col = "Sample_Size", vars = method_vars)
all_Increasing_res <- run_group(df_Increasing, success_col = "Increasing_n", denom_col = "Sample_Size", vars = all_substantive_vars)
```

```{r Increasing-mods-tables}
make_group_table(bg_res_Increasing, "Increasing Symptom Trajectory: Socio-Demographic Moderators")
make_group_table(tr_res_Increasing, "Increasing Symptom Trajectory: Trauma-Related Moderators")
make_group_table(md_res_Increasing, "Increasing Symptom Trajectory: Methodological Moderators")
```

Below is a summary of univariate models of all predictor variables for the Increasing Trajectory.

```{r Increasing-mods-tables-summary}
make_group_table(all_Increasing_res, "Increasing Symptom Trajectory: All Moderators")
```

## 🩸 High Symptom Trajectory

### Descriptives

We provide additional descriptives based on the High trajectory variable, using the full dataset.

```{r High-descriptives}
trajectory_descriptives(df_High, "High")
```

Below, we fit univariate models of predictors for the High Trajectory grouped by conceptual domains: Background, Trauma-related, and Methodological variables.
For continuous moderators, the model describes how the expected trajectory prevalence changes with a one-unit increase in the predictor. 
For categorical moderators, it compares the prevalence between the reference category and each alternative category.
Differences (Marginal Δ) are significant at α = .05. 

```{r High-mods-fitting}
# Run the three groups for this trajectory
bg_res_High <- run_group(df_High, success_col = "High_n", denom_col = "Sample_Size", vars = background_vars)
tr_res_High <- run_group(df_High, success_col = "High_n", denom_col = "Sample_Size", vars = trauma_vars)
md_res_High <- run_group(df_High, success_col = "High_n", denom_col = "Sample_Size", vars = method_vars)
all_High_res <- run_group(df_High, success_col = "High_n", denom_col = "Sample_Size", vars = all_substantive_vars)
```

```{r High-mods-tables}
make_group_table(bg_res_High, "High Symptom Trajectory: Socio-Demographic Moderators")
make_group_table(tr_res_High, "High Symptom Trajectory: Trauma-Related Moderators")
make_group_table(md_res_High, "High Symptom Trajectory: Methodological Moderators")
```

Below is a summary of univariate models of all predictor variables for the High Trajectory.

```{r High-mods-tables-summary}
make_group_table(all_High_res, "High Symptom Trajectory: All Moderators")
```

## 🌗 Moderate Symptom Trajectory

### Descriptives

We provide additional descriptives based on the Moderate trajectory variable, using the full dataset.

```{r Moderate-descriptives}
trajectory_descriptives(df_Moderate, "Moderate")
```

Below, we fit univariate models of predictors for the Moderate Trajectory grouped by conceptual domains: Background, Trauma-related, and Methodological variables.
For continuous moderators, the model describes how the expected trajectory prevalence changes with a one-unit increase in the predictor. 
For categorical moderators, it compares the prevalence between the reference category and each alternative category.
Differences (Marginal Δ) are significant at α = .05. 

```{r Moderate-mods-fitting}
# Run the three groups for this trajectory
bg_res_Moderate <- run_group(df_Moderate, success_col = "Moderate_n", denom_col = "Sample_Size", vars = background_vars)
tr_res_Moderate <- run_group(df_Moderate, success_col = "Moderate_n", denom_col = "Sample_Size", vars = trauma_vars)
md_res_Moderate <- run_group(df_Moderate, success_col = "Moderate_n", denom_col = "Sample_Size", vars = method_vars)
all_Moderate_res <- run_group(df_Moderate, success_col = "Moderate_n", denom_col = "Sample_Size", vars = all_substantive_vars)
```

```{r Moderate-mods-tables}
make_group_table(bg_res_Moderate, "Moderate Symptom Trajectory: Socio-Demographic Moderators")
make_group_table(tr_res_Moderate, "Moderate Symptom Trajectory: Trauma-Related Moderators")
make_group_table(md_res_Moderate, "Moderate Symptom Trajectory: Methodological Moderators")
```

Below is a summary of univariate models of all predictor variables for the Moderate Trajectory.

```{r Moderate-mods-tables-summary}
make_group_table(all_Moderate_res, "Moderate Symptom Trajectory: All Moderators")
```

# Model 2: Multivariate BRMS Models with Missing-Data Imputation

The second modelling stage estimates multivariate Bayesian regression models in which all moderators are entered simultaneously using a binomial likelihood. This approach allows us to assess the unique contribution of each predictor while adjusting for the influence of all other moderators in the model. By modelling all predictors jointly, these analyses move beyond simple associations and provide a more rigorous test of independent effects.

Continuous predictors are standardised to place effects on a comparable scale, and categorical predictors are entered as factors with defined reference categories. Missing values in continuous moderators are handled directly within the Bayesian framework using model-based imputation, allowing all available information to be retained rather than relying on listwise deletion. Prior to model fitting, correlation matrices are used to identify highly collinear predictors, with variables exceeding a correlation of .90 excluded to ensure model stability.

For each trajectory (Low, Decreasing, Increasing, High, and Moderate), three nested models are estimated: a background-only model, a model adding trauma-related moderators, and a full model including methodological variables. This hierarchical structure allows us to evaluate whether additional domains of moderators improve explanatory power beyond basic sample characteristics. Model results are summarised using posterior medians and 95% credible intervals, with convergence diagnostics used to verify estimation quality, and predictive performance compared across models using leave-one-out cross-validation.

## Correlation of Continous Predictors

Since the models incorporate multiple moderators simultaneously, it is important to understand how strongly continuous variables relate to one another. High correlations may signal redundancy or confounding. By computing and visualising the correlation matrix, we obtain an initial sense of the structure among moderators, helping us to anticipate potential collinearity and refine the interpretation of subsequent model coefficients.
Here, we first provide the correlation matrix of continuous variables of the whole dataset.

```{r correlation-table,  results='asis', message=FALSE}
# GT Correlation Table for Continuous Predictors
generate_correlation_gt <- function(df, 
                                    vars = NULL, 
                                    trajectory = NULL, 
                                    outcome = NULL,
                                    show_warnings = TRUE) {
  
  # Get continuous variables from MOD_SPECS
  cont_vars <- MOD_SPECS |> 
    purrr::keep(~ .x$type == "cont") |> 
    purrr::map_chr("var")
  
  # Keep only variables present in df
  cont_vars <- intersect(cont_vars, names(df))
  
  # Compute correlation matrix
  cor_mat <- round(
    cor(df[, cont_vars], use = "pairwise.complete.obs"),
    2
  )
  
  # Convert to long format
  cor_df <- cor_mat %>%
    as.data.frame() %>%
    tibble::rownames_to_column("Var1") %>%
    tidyr::pivot_longer(-Var1, names_to = "Var2", values_to = "Correlation")
  
  # Initialize exclusions
  exclude_vars <- character(0)
  
  # high correlation check
  if (show_warnings) {
    
    # Filter for high correlations and exclude self-pairs + duplicates
    high_cor_df <- cor_df %>%
      dplyr::filter(abs(Correlation) > 0.90 & Var1 != Var2) %>%
      dplyr::rowwise() %>%
      dplyr::filter(Var1 < Var2) %>%
      dplyr::ungroup()
    
    # If any high-correlation pairs exist:
    if (nrow(high_cor_df) > 0) {
      
      # Add pretty labels
      high_cor_df$Pretty1 <- MOD_LABELS[high_cor_df$Var1]
      high_cor_df$Pretty2 <- MOD_LABELS[high_cor_df$Var2]
      
      # Define exclusions for downstream analysis
      if (!is.null(outcome)) {
      
        # Correlation of each predictor with the outcome
        outcome_cor <- cor(
          df[, cont_vars, drop = FALSE],
          df[[outcome]],
          use = "pairwise.complete.obs"
        )
      
        outcome_cor <- abs(outcome_cor)
        names(outcome_cor) <- cont_vars  
        # match()
        high_cor_df <- high_cor_df %>%
          dplyr::arrange(desc(abs(Correlation)))
        
        dropped <- character(0)
        keep_results <- list()
        
        for (i in seq_len(nrow(high_cor_df))) {
          
          v1 <- high_cor_df$Var1[i]
          v2 <- high_cor_df$Var2[i]
          
          # Skip if either variable already dropped
          if (v1 %in% dropped || v2 %in% dropped) next
          
          cor1 <- outcome_cor[v1]
          cor2 <- outcome_cor[v2]
          
          if (cor1 >= cor2) {
            drop_var <- v1
            keep_var <- v2
          } else {
            drop_var <- v2
            keep_var <- v1
          }
          
          dropped <- c(dropped, drop_var)
          
          keep_results[[length(keep_results) + 1]] <- data.frame(
            Var1 = v1,
            Var2 = v2,
            drop_var = drop_var,
            keep_var = keep_var
          )
        }
        
        high_cor_df <- dplyr::bind_rows(keep_results)
        exclude_vars <- unique(high_cor_df$drop_var)
      }
            
      # warning: dropped variables
    if (length(exclude_vars) > 0) {
    
      dropped_pretty <- MOD_LABELS[exclude_vars]
    
      cat("<hr>")
      cat("<br><strong>Warning: Highly Correlated Variables Removed</strong><br>")
      cat(
        paste0(
          "The following continuous predictors were dropped due to high correlation ",
          "(|r| > 0.90) based on their weaker association with the outcome ",
          "for the <em>", trajectory, "</em> trajectory:<br><br>",
          "<em>",
          paste(dropped_pretty, collapse = ", "),
          "</em><br><br>"
        )
      )
    
    }
      
    } else {
      cat("\n--- No variable pairs found with correlation > 0.90. ---\n\n")
    }
  }
  
  # Add pretty labels for GT table
  cor_df$Pretty1 <- MOD_LABELS[cor_df$Var1]
  cor_df$Pretty2 <- MOD_LABELS[cor_df$Var2]
  
  # 6. Pivot to wide format
  cor_wide <- cor_df %>%
    dplyr::select(Pretty1, Pretty2, Correlation) %>%
    tidyr::pivot_wider(
      names_from = Pretty2,
      values_from = Correlation
    ) %>%
    dplyr::rename(Variable = Pretty1)
  
  # Build GT table
  cor_gt <- cor_wide %>%
    gt::gt(rowname_col = "Variable") %>%
    gt::tab_header(
      title = paste0(trajectory, " Trajectory Correlation Matrix: Continuous Variables"),
      subtitle = "Pearson correlations (pairwise complete observations)"
    ) %>%
    gt::fmt_number(columns = gt::everything(), decimals = 2) %>%
    gt::data_color(
      columns = gt::everything(),
      fn = scales::col_numeric(
        palette = c("red", "white", "red"),
        domain = c(-1, 1)
      )
    ) %>%
    gt::tab_options(
      table.font.size = 12,
      data_row.padding = gt::px(3),
      column_labels.font.weight = "bold",
      table.font.names = "Times New Roman"
    )
}

correlations_df <- generate_correlation_gt(df_moderation, show_warnings = FALSE)
correlations_df
```

Due to high collinearity, the variable "Time span of the study (months)" (a.k.a. TP1-TPX) will be excluded from all analyses for this Bayesian model. Trajectory-specific correlation matrices are provided in the corresponding sections of this model and variables that have a correlation index of > .90 are excluded from the analyses. 
Further, to account for data sparcity and high overlap, some categorical variables will be combined/omitted from the analysis. Due to high overlap between the variable "Military" and the level "Combat" of the variable "Trauma type", the variable "Military" will also be excluded from all analyses for this Bayesian model.

## Data Preperation for BRMS

Before fitting multivariate models, the data are prepared for Bayesian estimation and missing-data imputation. Continuous moderators are standardised to place them on a comparable scale, and categorical variables are recoded to ensure proper factor structure. This preparation ensures stable estimation and interpretable coefficients in the multivariate analyses.

```{r BRMS-data-prep}
# Differentiate between categorical and continuous variables
cont_vars <- purrr::map_chr(
  purrr::keep(MOD_SPECS, ~ .x$type == "cont"),
  "var"
)

cat_vars <- purrr::map_chr(
  purrr::keep(MOD_SPECS, ~ .x$type == "cat"),
  "var"
)

# Function to apply model specifications to a data frame
apply_mod_specs <- function(df, mod_specs) {
  
  # Iterate over each specification
  for (spec in mod_specs) {
    
    var  <- spec$var
    type <- spec$type
    
    # Skip if variable is not in the data frame
    if (!var %in% names(df)) next
    
    
    # Handle continuous variables
    if (identical(type, "cont")) {
      
      x_raw <- df[[var]]
      # Coerce to numeric
      x_num <- suppressWarnings(as.numeric(x_raw))
      
      # Warning check
      if (all(is.na(x_num)) && any(!is.na(x_raw))) {
        warning("Variable '", var, "' became all NA when coerced to numeric.")
      }
      
      df[[var]] <- x_num
      next
    }
    
    
    # Handle categorical variables
    if (identical(type, "cat")) {
      
      # Clean string values
      x <- df[[var]] |>
        as.character() |>
        trimws() |>
        stringr::str_squish() 

      
      # Apply level mapping
      if (!is.null(spec$levels)) {
        lvl_map  <- spec$levels
        raw_keys <- names(lvl_map)
        
        # Apply mapping
        matched  <- x %in% raw_keys
        x[matched] <- lvl_map[x[matched]]
        
        # Report unmatched raw values
        unmatched_raw <- sort(unique(x_raw[!(x_raw %in% raw_keys) & !is.na(x_raw)]))
        if (length(unmatched_raw) > 0L) {
          message(
            "apply_mod_specs: in variable '", var,
            "', some raw values were not mapped by 'levels': ",
            paste(unmatched_raw, collapse = ", ")
          )
        }
      }
      
      
      # Filter to keep only specified levels/labels
      if (!is.null(spec$keep)) {
        if (!is.null(spec$levels)) {
          # Get the mapped labels
          keep_labels <- unname(spec$levels[spec$keep])
        } else {
          # Use 'keep' values directly
          keep_labels <- spec$keep
        }
        
        # Set non-kept values to NA
        x[!(x %in% keep_labels)] <- NA
      }
      
      # Handle missing values (NA) by mapping them to the na_level string
      if (!is.null(spec$na_level)) {
        x[is.na(x)] <- spec$na_level
      }

      # Convert to factor
      x <- factor(x)
      levels(x) <- trimws(levels(x))
      
      # Set reference level
      if (!is.null(spec$ref)) {
        ref_label <- spec$ref
        
        # Use mapped label for reference if available
        if (!is.null(spec$levels) && spec$ref %in% names(spec$levels)) {
          ref_label <- spec$levels[[spec$ref]]
        }
        
        # Relevel
        if (ref_label %in% levels(x)) {
          x <- stats::relevel(x, ref = ref_label)
        } else {
          warning(
            "Reference level '", spec$ref,
            "' (mapped to '", ref_label,
            "') not found in levels of '", var, "'."
          )
        }
      }
      
      df[[var]] <- x
    }
  }
  
  df
}

# Function to prepare data for brms (apply specs and scale/factor)
prepare_data_brms <- function(df, mod_specs = MOD_SPECS, exclude = NULL) {
  
  # Remove excluded variables from df entirely
  if (!is.null(exclude)) {
    df <- df %>% dplyr::select(-dplyr::all_of(exclude))
  }
  
  # Extract names of continuous variables
  cont_vars <- purrr::map_chr(
    purrr::keep(mod_specs, ~ .x$type == "cont"),
    "var"
  )
  
  # Extract names of categorical variables
  cat_vars <- purrr::map_chr(
    purrr::keep(mod_specs, ~ .x$type == "cat"),
    "var"
  )
  
  # Identify variables present in df
  cont_vars_in_df <- intersect(cont_vars, names(df))
  cat_vars_in_df  <- intersect(cat_vars,  names(df))
  
  # Apply all variable specifications (recode, filter, set ref level)
  df <- apply_mod_specs(df, mod_specs)
  
  # Scale continuous variables
  if (length(cont_vars_in_df) > 0) {
    df <- df %>%
      dplyr::mutate(
        dplyr::across(
          dplyr::all_of(cont_vars_in_df),
          ~ as.numeric(scale(suppressWarnings(as.numeric(.))))
        )
      )
  }
  
  #  Convert binary factors to 0/1 numeric
  binary_vars_to_encode <- intersect(get_binary_cat_vars(df, cat_vars_in_df), names(df))
  
  if (length(binary_vars_to_encode) > 0) {
    df <- df %>%
      dplyr::mutate(
        dplyr::across(
          dplyr::all_of(binary_vars_to_encode),
          ~ factor_to_01(.) # Converts two-level factor to 0/1 numeric
        )
      )
  }
  # Ensure all other categorical variables are factors
  non_binary_cat_vars <- setdiff(cat_vars_in_df, binary_vars_to_encode)
  if (length(non_binary_cat_vars) > 0) {
    df <- df %>%
      dplyr::mutate(
        dplyr::across(
          dplyr::all_of(non_binary_cat_vars),
          ~ factor(.)
        )
      )
  }
  
  # Drop unused levels only for the remaining factors (non-binary)
  df <- droplevels(df)
  
  return(df)
}
```

## BMRS setup

The Bayesian models are implemented using the brms package, which provides flexible model specification and seamless integration of missing-data imputation via the mi() syntax. The chosen sampling settings balance computational feasibility with the need for stable posterior estimates.

```{r BRMS-setup}
# Use all cores for parallel chains
options(mc.cores = parallel::detectCores())
brms_args <- list(
  family  = binomial(link = "logit"),
  chains  = 4,
  iter    = 2000,
  warmup  = 300,
  cores   = 4,
  seed    = 1,
  backend = "cmdstanr",
  refresh = 0,
  control = list(
    adapt_delta   = 0.99
  ),
  stan_model_args = list(
  cpp_options = list(CXXFLAGS = "-O3")
  )
)
```

## Portable Functions for Model Fitting

This section defines functions that construct the multivariate models automatically for each trajectory. Importantly, missing predictors are estimated through univariate imputation. This approach avoids the loss of information that would occur with listwise deletion and enables the inclusion of a broad predictor set even when individual studies did not report all moderators.

### Fitting the Multivariate Bayesian Models

The multivariate models move beyond simple associations and estimate the unique contribution of each predictor while adjusting for all others in the model. We estimate three nested models per trajectory:

M1: Background-variables-only model

M2: Background + Trauma-related variables model

M3: Full model (Background + Trauma + Methodological variables)

Comparing these models helps determine whether trauma-related or methodological factors contribute additional explanatory value beyond basic sample characteristics.
These groups are widely the same across trajectories, but may differ slightly based on exclusions due to collinearity.

```{r BRMS-model-fitting-functions}
get_binary_cat_vars <- function(df, cat_vars_all) {
  binary_vars <- c()
  for (v in intersect(cat_vars_all, names(df))) {
    x_non_na <- df[[v]][!is.na(df[[v]])]
    if (is.factor(x_non_na) && nlevels(x_non_na) == 2) {
      binary_vars <- c(binary_vars, v)
    }
  }
  return(binary_vars)
}
factor_to_01 <- function(fct_var) {
  if (!is.factor(fct_var) || nlevels(fct_var) != 2) {
    stop("Input must be a two-level factor.")
  }
  # Converts factor levels (1, 2) to (0, 1), where the reference level (1) becomes 0.
  return(as.numeric(fct_var) - 1)
}

## Build predictor families
build_predictor_families <- function(df, cont_vars, cat_vars) {
  
  fams <- setNames(
    rep(list(gaussian()), length(cont_vars)),
    cont_vars
  )
  
  for (v in cat_vars) {
    if (!v %in% names(df)) next
    
    x <- factor(df[[v]])
    nlev <- nlevels(x)
    
    if (nlev < 2) {
      stop("Variable ", v, " has <2 levels after cleaning.")
    }
    
    # Use categorical() for ALL categorical predictors (2+ levels)
    fams[[v]] <- categorical()
  }
  
  fams
}


## univariate imputation formula 
fit_brms_models_mi <- function(df,
                               success_col_name,
                               model_vars,
                               cont_vars,
                               brms_args) {
  if (!success_col_name %in% names(df)) {
    stop("Outcome variable '", success_col_name, "' not found in df.")
  }
  
  model_vars <- intersect(model_vars, names(df))
  model_vars <- setdiff(model_vars, success_col_name)
  
  if (length(model_vars) == 0L) {
    stop("No valid predictor variables found in 'model_vars'.")
  }
  
  # vars that have missingness
  vars_with_na_raw <- model_vars[
    sapply(df[, model_vars, drop = FALSE], function(x) any(is.na(x)))
  ]
  
  # Split missing variables by imputed type (Numeric vs. Factor)
  # Numeric includes original continuous and the new 0/1 binary variables (Gaussian family)
  vars_numeric_na <- vars_with_na_raw[
    sapply(df[, vars_with_na_raw, drop = FALSE], is.numeric)
  ]
  
  # Factor variables (Multi-level categorical, also imputed with Gaussian approximation)
  vars_factor_na <- vars_with_na_raw[
    sapply(df[, vars_with_na_raw, drop = FALSE], is.factor)
  ]
  
  vars_with_na <- c(vars_numeric_na, vars_factor_na)
  
  # Predictors without NAs (no imputation needed)
  vars_without_na <- setdiff(model_vars, vars_with_na)
  
  # Build RHS terms for the main model (Outcome ~ Predictors)
  rhs_terms <- c(
    if (length(vars_with_na)) paste0("mi(", vars_with_na, ")"),
    setdiff(vars_without_na, "Study") 
  )
  
  # Main formula for the outcome (binomial GLMM)
  main_formula <- bf(
    stats::as.formula(
      paste0(
        success_col_name,
        " | trials(Sample_Size) ~ ",
        paste(rhs_terms, collapse = " + "),
        " + (1|Study)" 
      )
    ),
    family = binomial(link = "logit")
  )
  
  # Case 1: no MI → simple model
  if (length(vars_with_na) == 0L) {
    default_priors <- brms::get_prior(formula = main_formula, data = df) 
    fit <- do.call(
      brms::brm,
      c(
        list(
          formula = main_formula,
          data    = df, 
          prior   = default_priors
        ),
        brms_args
      )
    )
    return(fit)
  }
  
 # Case 2: MI for Numeric (Continuous & 0/1) and Factor (Multi-level) vars
  
  # Imputation formulas for Numeric variables (Gaussian family)
  mi_formulas_numeric <- lapply(vars_numeric_na, function(v) {
    brms::bf(stats::as.formula(paste0(v, " | mi() ~ 1")), family = stats::gaussian()) 
  })
  
  # Imputation formulas for Factor variables (Gaussian approximation for simplicity/compatibility)
  mi_formulas_factor <- lapply(vars_factor_na, function(v) {
    message(paste("Warning: Imputing multi-level factor", v, "using Gaussian approximation."))
    brms::bf(stats::as.formula(paste0(v, " | mi() ~ 1")), family = stats::gaussian()) 
  })
  
  mi_formulas <- c(mi_formulas_numeric, mi_formulas_factor)
  
  full_formula <- Reduce(`+`, c(list(main_formula), mi_formulas))
  
  default_priors <- brms::get_prior(formula = full_formula, data = df) 
  
  fit <- do.call(
    brms::brm,
    c(
      list(
        formula = full_formula,
        data    = df, 
        prior   = default_priors
      ),
      brms_args
    )
  )
  
  return(fit)
}

# looping logic for running all models for one trajectory
fit_brms_trajectory_models <- function(df,
                                       success_col_name,
                                       cont_vars,
                                       cat_vars,
                                       mod_specs,
                                       background_vars,
                                       trauma_vars,
                                       method_vars,
                                       brms_args,
                                       exclude_vars = NULL) {
  
  message("Starting fitting for trajectory with outcome: ", success_col_name)
  
  # multicollinearity filtering step
  
  if (!is.null(exclude_vars) && length(exclude_vars) > 0) {
    message("\nExcluding variables due to high correlations:\n",
            paste(" -", exclude_vars, collapse = "\n"))
  }
  
  cont_vars       <- setdiff(cont_vars, exclude_vars)
  cat_vars        <- setdiff(cat_vars, exclude_vars)
  background_vars <- setdiff(background_vars, exclude_vars)
  trauma_vars     <- setdiff(trauma_vars, exclude_vars)
  method_vars     <- setdiff(method_vars, exclude_vars)

  # Model variable setups
  
  model_list <- list(
    model_1_bg        = background_vars,
    model_2_bg_trauma = c(background_vars, trauma_vars),
    model_3_full      = c(background_vars, trauma_vars, method_vars)
  )

  # Fit the models
  
  purrr::map(
    model_list,
    ~ fit_brms_models_mi(
      df               = df,
      success_col_name = success_col_name,
      model_vars       = .x,
      cont_vars        = cont_vars,
      brms_args        = brms_args
    )
  )
}
```

### Summary Tables

We translate posterior distributions into clear summaries that report the median effect size and 95% credible interval for each predictor. 
Predictors whose credible intervals exclude zero are flagged as potentially important. Presenting all three models side-by-side allows readers to see how effects evolve as more predictors are added.

Model diagnostics summarise convergence metrics such as R-hat and effective sample sizes. These indicators ensure that posterior estimates are reliable and that inferences from the multivariate models can be trusted.

Finally, the multivariate models are compared using leave-one-out cross-validation (LOO-CV). This provides an estimate of the relative predictive accuracy of the three model specifications and helps assess whether additional moderator domains improve the overall model fit for each trajectory. LOO-CV ranks models by how well they predict unseen data using ELPD, where higher ELPD values indicate better predictive accuracy. The best model has ΔELPD = 0, and other models have more negative ΔELPD values the worse they perform, with the SE showing how confidently we can distinguish those differences. Model weights summarize each model’s relative predictive support, where higher weights signal a model that is more likely to generalize well to new data.

```{r BRMS-gt-table}
# Predictor Summary Table Function
generate_predictor_table_gt <- function(fitted_models,
                                        success_col_name,
                                        trajectory_name,
                                        mod_specs = MOD_SPECS) {
  
  var_names  <- purrr::map_chr(mod_specs, "var") # Extract variable names from specifications
  var_labels <- purrr::map_chr(mod_specs, "label") |> # Extract variable labels
    stats::setNames(var_names) # Name labels with variable names
  
  find_base_var <- function(t0) {

    # Order candidate variable names from longest to shortest
    hits <- var_names[order(nchar(var_names), decreasing = TRUE)]
  
    # Keep only those whose names appear in the term
    hit <- purrr::keep(hits, ~ grepl(.x, t0))
  
    # If nothing matches, return original term
    if (length(hit) == 0) return(t0)
  
    # Return the first (i.e., longest match)
    hit[[1]]
  }

  all_coefs <- purrr::map_dfr(names(fitted_models), function(model_name) { # Iterate over fitted models
    coefs <- try(brms::fixef(fitted_models[[model_name]], summary = TRUE), silent = TRUE) # Get fixed effects summary
    if (inherits(coefs, "try-error") || length(coefs) == 0) return(NULL) # Skip if error or no coefficients
    
    as.data.frame(coefs) %>%
      tibble::rownames_to_column(var = "Term") %>% # Convert row names (terms) to a column
      dplyr::rename( # Rename columns for clarity
        Median   = Estimate,
        CI_Lower = Q2.5,
        CI_Upper = Q97.5
      ) %>%
      dplyr::mutate(Model = model_name) # Add model name column
  })
  
  coefficient_data <- all_coefs %>%
    dplyr::filter(!grepl("Intercept", Term)) %>% # Remove intercept terms
    dplyr::mutate(
      Term = gsub("^Lown_|^Decreasingn_|^Increasingn_|^Highn_|^Moderaten_", "", Term), # Clean up term names (remove prefixes)
      Term = sub("^mi_", "", Term),
      Term = sub("^mi", "", Term)
 # Remove 'mi' (multiple imputation indicator)
    ) %>%
    dplyr::rowwise() %>% # Operate row by row
    dplyr::mutate(
      BaseVar = find_base_var(Term), # Find the base variable name
      Level   = trimws(sub(paste0("^", BaseVar), "", Term)), # Extract the level/suffix of the variable
      VarLabel = ifelse(BaseVar %in% names(var_labels), # Get the descriptive label for the base variable
                        var_labels[BaseVar],
                        BaseVar),
      Predictor = ifelse( # Create the final predictor label
        Level == "" | is.na(Level), # If no level suffix, use the base label (e.g., for continuous variables)
        VarLabel,
        # If a level exists, combine base label and level in parentheses
        paste0(VarLabel, " (", Level, ")") 
      )
    ) %>%
    dplyr::ungroup() %>%
    dplyr::filter(
      !(grepl("Trauma type", Predictor) & grepl("(Other)", Predictor))
    ) %>%
    dplyr::mutate(
      Effect = paste0( # Format effect size and 95% CI
        sprintf("%.2f", Median), " [",
        sprintf("%.2f", CI_Lower), ", ",
        sprintf("%.2f", CI_Upper), "]"
      ),
      Important = ifelse(CI_Lower > 0 | CI_Upper < 0, "Yes", "No"), # Check if CI excludes zero
      Model = dplyr::recode( # Recode model names for display
        Model,
        "model_1_bg"        = "M1 (Background)",
        "model_2_bg_trauma" = "M2 (Background + Trauma)",
        "model_3_full"      = "M3 (Background + Trauma + Methodological)"
      )
    ) %>%
    dplyr::select(Predictor, Model, Effect, Important) # Select final columns
  
  coefficient_wide <- coefficient_data %>%
    tidyr::pivot_wider( # Reshape data from long to wide format
      names_from  = Model, # Columns created based on Model
      values_from = c(Effect, Important), # Values to spread across new columns
      names_glue  = "{.value}_{Model}" # Naming convention for new columns
    )
  
  effect_cols     <- grep("^Effect_", names(coefficient_wide), value = TRUE) # Identify Effect columns
  importance_cols <- grep("^Important_", names(coefficient_wide), value = TRUE) # Identify Important columns
  
  effect_labels <- stats::setNames( # Create labels for Effect columns
    gsub("Effect_", "Effect [95% CI] – ", effect_cols),
    effect_cols
  )
  
  importance_labels <- stats::setNames( # Create labels for Important columns
    gsub("Important_", "Important? – ", importance_cols),
    importance_cols
  )
  
  coefficient_wide %>%
    gt::gt(rowname_col = "Predictor") %>% # Initialize gt table, setting Predictor as row names
    gt::tab_header( # Add table title and subtitle
      title    = gt::md(glue::glue("**Predictor Summary: {trajectory_name} Trajectory**")),
      subtitle = "Median, 95% Credible Interval, and Predictor Importance"
    ) %>%
    gt::cols_label( # Apply custom column labels
      !!!effect_labels,
      !!!importance_labels
    ) %>%
    gt::data_color( # Color cells based on 'Important' status
      columns = dplyr::all_of(importance_cols),
      fn  = function(x) dplyr::case_when(
        x == "Yes" ~ "#d4edda", # Light green for 'Yes'
        x == "No"  ~ "lightgray",   # Light gray for 'No'
        TRUE       ~ "white"
      )
    ) %>%
    gt::tab_options(
      table.font.size = 12,
      data_row.padding = gt::px(3),
      column_labels.font.weight = "bold",
      table.font.names = "Times New Roman"
    )
}

# Diagnostic Table Function
diagnostic_gt_brms <- function(model,
                               model_name = NULL,
                               trajectory_name = NULL,
                               include = c("fixed", "sigma"),
                               title = "Model Diagnostics",
                               subtitle = NULL,
                               mod_labels = MOD_LABELS) {

  s <- summary(model)
  parts <- list()
  # Canonicalize coefficient names
  canonical_term <- function(term) {
    term <- gsub("`", "", term)
    term <- sub("^[^_]+_", "", term)   # drop outcome prefix
    term <- sub("^mi", "", term)       # drop mi prefix
    term
  }
  # Detect categorical moderator (longest name wins)
  cat_vars <- names(CAT_SPECS)
  cat_vars <- cat_vars[order(nchar(cat_vars), decreasing = TRUE)]
  detect_cat_var <- function(term) {
    tt <- canonical_term(term)
    if (tt %in% cat_vars) return(tt)
    for (v in cat_vars) {
      if (startsWith(tt, v)) return(v)
    }
    NA_character_
  }
  # Pretty predictor labels
  pretty_from_mod <- function(term) {
    tt <- canonical_term(term)
    if (tt %in% names(mod_labels)) return(mod_labels[[tt]])
    v <- detect_cat_var(term)
    if (!is.na(v) && v %in% names(mod_labels)) return(mod_labels[[v]])
    term
  }
  
  pretty_sigma <- function(term) {
    # exact sigma
    if (term == "sigma")
      return("Residual SD")
  
    if (!startsWith(term, "sigma_"))
      return(term)
  
    raw <- sub("^sigma_", "", term)
  
    # direct match
    if (raw %in% names(MOD_LABELS)) {
      return(paste0("Residual SD: ", MOD_LABELS[[raw]]))
    }
  
    # underscore-insensitive match
    raw_clean  <- gsub("_", "", raw)
    keys_clean <- gsub("_", "", names(MOD_LABELS))
  
    idx <- match(raw_clean, keys_clean)
  
    if (!is.na(idx)) {
      return(paste0(
        "Residual SD: ",
        MOD_LABELS[[names(MOD_LABELS)[idx]]]
      ))
    }
  
    paste0("Residual SD: ", raw)
  }
  # Contrast extractor
  extract_contrast_info <- function(term) {
    var <- detect_cat_var(term)
    if (is.na(var)) return("")
    spec <- CAT_SPECS[[var]]
    lvls <- LEVEL_MAP |> dplyr::filter(moderator == var)
    if (nrow(lvls) == 0) return("")
    # reference
    ref_code <- if (!is.null(spec$ref)) spec$ref else lvls$orig[1]
    ref_pretty <- lvls |>
      dplyr::filter(orig == ref_code) |>
      dplyr::pull(pretty)
    ref_pretty <- if (length(ref_pretty)) ref_pretty[1] else ref_code
    tt <- canonical_term(term)
    # CASE 1: binary categorical converted to 0/1
    if (tt == var) {
      if (!is.null(spec$keep) && length(spec$keep) >= 2) {
        other_code <- spec$keep[spec$keep != ref_code][1]
      } else {
        other_code <- lvls$orig[lvls$orig != ref_code][1]
      }
      other_pretty <- lvls |>
        dplyr::filter(orig == other_code) |>
        dplyr::pull(pretty)
      other_pretty <- if (length(other_pretty)) other_pretty[1] else other_code
      return(paste0(ref_pretty, " → ", other_pretty))
    }
    # CASE 2: multi-level categorical (suffix present)
    token <- sub(paste0("^", var), "", tt)
    if (nzchar(token)) {
      lev_pretty <- lvls |>
        dplyr::filter(orig == token) |>
        dplyr::pull(pretty)
      lev_pretty <- if (length(lev_pretty)) lev_pretty[1] else token
      return(paste0(ref_pretty, " → ", lev_pretty))
    }
    ""
  }
  term_is_cat <- function(term) {
    !is.na(detect_cat_var(term))
  }
  # Fixed effects
  if ("fixed" %in% include) {
    fx <- as.data.frame(s$fixed)
    fx$term <- rownames(fx)
    fx <- fx |>
      dplyr::filter(!grepl("Intercept", term, ignore.case = TRUE)) |>
      dplyr::mutate(
        Predictor = purrr::map_chr(term, pretty_from_mod),
        Contrast  = purrr::map_chr(term, extract_contrast_info),
        type      = ifelse(vapply(term, term_is_cat, logical(1)), "cat", "cont")
      ) |>
      dplyr::arrange(desc(type))
    parts$fixed <- fx |>
      dplyr::select(Predictor, Contrast,
                    Estimate, Est.Error, Rhat, Bulk_ESS, Tail_ESS)
  }
  # Sigma parameters
  if ("sigma" %in% include && !is.null(s$spec_pars)) {
  
    sg <- as.data.frame(s$spec_pars)
    sg$term <- rownames(sg)
  
    sg <- sg |>
      dplyr::filter(!grepl("Intercept", term, ignore.case = TRUE)) |>
      dplyr::mutate(
        Predictor = purrr::map_chr(term, pretty_sigma),
        Contrast  = ""
      )
  
    parts$sigma <- sg |>
      dplyr::select(Predictor, Contrast,
                    Estimate, Est.Error, Rhat, Bulk_ESS, Tail_ESS)
  }
  # Combine + GT
  tab <- dplyr::bind_rows(parts, .id = "component")
  gt::gt(tab, groupname_col = "component") |>
    gt::tab_header(
      title = paste0(
        title,
        if (!is.null(model_name)) paste0(" — ", model_name),
        if (!is.null(trajectory_name)) paste0(" (", trajectory_name, ")")
      ),
      subtitle = subtitle
    ) |>
    gt::fmt_number(
      columns = c(Estimate, Est.Error, Rhat, Bulk_ESS, Tail_ESS),
      decimals = 3
    ) |>
    gt::cols_label(
      Predictor = "Predictor",
      Contrast  = "Contrast (Ref → Level)",
      Estimate  = "Estimate",
      Est.Error = "Std. Error",
      Rhat      = "Rhat",
      Bulk_ESS  = "Bulk ESS",
      Tail_ESS  = "Tail ESS"
    ) |>
    gt::opt_row_striping() |>
    gt::tab_style(
      style = gt::cell_text(weight = "bold"),
      locations = gt::cells_column_labels()
    ) |>
    gt::tab_options(
      table.font.size = 12,
      table.font.names = "Times New Roman"
    )
}

# Model Comparison Function
loo_gt_table <- function(
  model_list,
  model_names = NULL,
  response = NULL,
  table_title = "Model Comparison (LOO-CV)",
  add_weights = TRUE,              # compute stacking / Pseudo-BMA weights
  weight_method = c("stacking", "pseudobma")
) {
  
  weight_method <- match.arg(weight_method)

  # Model names
  if (is.null(model_names)) {
    model_names <- paste0("Model ", seq_along(model_list))
  }
  stopifnot(length(model_names) == length(model_list))

  # Compute LOO for each model
  loo_list <- lapply(
    model_list,
    function(m) {
      if (!is.null(response)) loo(log_lik(m, resp = response))
      else loo(log_lik(m))
    }
  )

  # Compare LOO models
  loo_comp <- do.call(loo_compare, loo_list)

  # Prepare the df
  df <- loo_comp %>%
    as.data.frame() %>%
    rownames_to_column("Model") %>%
    rename(
      # Only columns present in loo_compare:
      ELPD_diff = elpd_diff,
      SE = se_diff
    ) %>%
    mutate(
      ELPD_diff = round(ELPD_diff, 3),
      SE = round(SE, 3)
    ) %>%
    select(Model, ELPD_diff, SE)

  # Add model weights 
  if (add_weights) {
    if (weight_method == "stacking") {
      w <- loo_model_weights(loo_list, method = "stacking")
    } else if (weight_method == "pseudobma") {
      w <- loo_model_weights(loo_list, method = "pseudobma")
    }

    df$Weight <- round(as.numeric(w[match(df$Model, names(w))]), 3)
  }

  # Build gt table 
  gt_tbl <- df %>%
    gt() %>%
    tab_header(
      title = md(paste0("**", table_title, "**")),
      subtitle =
        "Higher ELPD indicates better out-of-sample predictive performance (with model averaging weights)"
    ) %>%
    cols_label(
      Model = "Model",
      ELPD_diff = "Δ ELPD (relative to best model)",
      SE = "SE(Δ)",
      Weight = "Model Weight"
    ) %>%
    fmt_number(
      columns = c(ELPD_diff, SE, dplyr::any_of("Weight")),
      decimals = 3
    ) %>%
    # Highlight best model: highest ELPD (i.e., Δ = 0)
    tab_style(
      style = list(
        cell_fill(color = "#f0f0f0"),
        cell_text(weight = "bold")
      ),
      locations = cells_body(rows = ELPD_diff == 0)
    ) %>%
    opt_row_striping() %>%
    opt_table_outline() %>%
    tab_options(
      table.font.size = 14,
      data_row.padding = px(6),
      heading.background.color = "#e8e8ef"
    )

  return(gt_tbl)
}
```

## 🚀 Low Symptom Trajectory

### Data Preparation and Correlation of Continuous Variables

Before fitting the multivariate models, data are prepared for Bayesian estimation through standardising continuous moderators and recoding categorical variables, ensuring stable estimation and interpretable coefficients.
A correlation matrix for the Low trajectory is used to detect highly collinear variables, with those exceeding a .90 correlation being excluded from analyses.

```{r BRMS-correlations-low, results='asis', message=FALSE}
# Prepare the data (cleaning, standardizing, recoding)
df_Low_clean <- prepare_data_brms(df_Low, exclude = c("TP1_TPX", "Military"))

# Determine Collinearity
correlations_Low <- generate_correlation_gt(df = df_Low_clean, trajectory = "Low", outcome = "Low_n" )
correlations_Low

exclude_vars_Low <- correlations_Low$exclude_vars
```

### Model Fits and Diagnostics

The analysis of the Low Trajectory multivariate models distils posterior distributions by reporting each predictor’s median effect size and 95% credible interval, highlighting predictors whose intervals do not include zero. Displaying the three models together shows how effects change as additional predictors are introduced. Model diagnostics, including R-hat and effective sample sizes, confirm good convergence and lend confidence to the reliability of the posterior estimates and resulting inferences.

```{r BRMS-fitting-low, results='asis', message=FALSE}
# Fit Models for Low Trajectory
models_Low <- fit_brms_trajectory_models(
  df               = df_Low_clean,
  success_col_name = "Low_n",
  cont_vars        = cont_vars,
  cat_vars         = cat_vars,
  mod_specs        = MOD_SPECS,
  background_vars  = background_vars,
  trauma_vars      = trauma_vars,
  method_vars      = method_vars,
  brms_args        = brms_args,
  exclude_vars     = exclude_vars_Low
)

# Produce predictor summary table
table_low <- generate_predictor_table_gt(
  fitted_models = models_Low,
  trajectory_name = "Low"
)
table_low

# Diagnostics for each submodel
diag_low_m1 <- diagnostic_gt_brms(model = models_Low$model_1_bg,
                   model_name = "M1: Background",
                   trajectory_name = "Low")

diag_low_m2 <- diagnostic_gt_brms(model = models_Low$model_2_bg_trauma,
                   model_name = "M2: Background + Trauma",
                   trajectory_name = "Low")

diag_low_m3 <- diagnostic_gt_brms(model = models_Low$model_3_full,
                   model_name = "M3: Background + Trauma + Methodological",
                   trajectory_name = "Low")

diag_low_m1
diag_low_m2
diag_low_m3
```

### Model Comparison

We compare the multivariate models of the Low trajectory using leave-one-out cross-validation (LOO-CV) to evaluate whether additional moderator domains improve predictive accuracy, ranking models by ELPD (with higher values indicating better performance), interpreting ΔELPD and its SE to gauge relative fit, and using optional model weights to summarize each model’s likelihood of generalizing well to new data.

```{r BRMS-model-comparison-low, results='asis', message=FALSE}
# Compare models using LOO-CV
# 1. Add LOO to all models in the list
models_Low <- purrr::map(models_Low, function(m) {
  add_criterion(m, "loo", resp = "Lown", moment_match = TRUE)
})
# 2. Compare them using the 'invoke' or 'do.call' method
comparison <- loo_compare(models_Low)
# 3. Print the results
print(comparison, simplify = FALSE)# Compare models using LOO-CV

gt_results_Low <- loo_gt_table(
  model_list = list(
    models_Low$model_1_bg,
    models_Low$model_2_bg_trauma,
    models_Low$model_3_full
  ),
  model_names = c( "M1 (Background)", "M2 (Background + Trauma)", "M3 (Background + Trauma + Methodological)"),
  response = "Lown"
)

gt_results_Low
```

## 🌿 Decreasing Symptom Trajectory

### Data Preparation and Correlation of Continuous Variables

Before fitting the multivariate models, data are prepared for Bayesian estimation through standardising continuous moderators and recoding categorical variables, ensuring stable estimation and interpretable coefficients.
A correlation matrix for the Decreasing trajectory is used to detect highly collinear variables, with those exceeding a .90 correlation being excluded from analyses.

```{r BRMS-correlations-Decreasing, results='asis', message=FALSE}
# Prepare the data (cleaning, standardizing, recoding)
df_Decreasing_clean <- prepare_data_brms(df_Decreasing, exclude = c("TP1_TPX", "Military"))

# Determine Collinearity
correlations_Decreasing <- generate_correlation_gt(df = df_Decreasing_clean, trajectory = "Decreasing", outcome = "Decreasing_n")
correlations_Decreasing

exclude_vars_Decreasing <- correlations_Decreasing$exclude_vars
```

### Model Fits and Diagnostics

The analysis of the Decreasing Trajectory multivariate models distils posterior distributions by reporting each predictor’s median effect size and 95% credible interval, highlighting predictors whose intervals do not include zero. Displaying the three models together shows how effects change as additional predictors are introduced. Model diagnostics, including R-hat and effective sample sizes, confirm good convergence and lend confidence to the reliability of the posterior estimates and resulting inferences.

```{r BRMS-fitting-Decreasing, results='asis', message=FALSE}
# Fit Models for Decreasing Trajectory
models_Decreasing <- fit_brms_trajectory_models(
  df               = df_Decreasing_clean,
  success_col_name = "Decreasing_n",
  cont_vars        = cont_vars,
  cat_vars         = cat_vars,
  mod_specs        = MOD_SPECS,
  background_vars  = background_vars,
  trauma_vars      = trauma_vars,
  method_vars      = method_vars,
  brms_args        = brms_args,
  exclude_vars     = exclude_vars_Decreasing
)

# Produce predictor summary table
table_Decreasing <- generate_predictor_table_gt(
  fitted_models = models_Decreasing,
  trajectory_name = "Decreasing"
)
table_Decreasing

# Diagnostics for each submodel
diag_Decreasing_m1 <- diagnostic_gt_brms(model = models_Decreasing$model_1_bg,
                                  model_name = "M1: Background",
                                  trajectory_name = "Decreasing")

diag_Decreasing_m2 <- diagnostic_gt_brms(model = models_Decreasing$model_2_bg_trauma,
                                  model_name = "M2: Background + Trauma",
                                  trajectory_name = "Decreasing")

diag_Decreasing_m3 <- diagnostic_gt_brms(model = models_Decreasing$model_3_full,
                                  model_name = "M3: Background + Trauma + Methodological",
                                  trajectory_name = "Decreasing")

diag_Decreasing_m1
diag_Decreasing_m2
diag_Decreasing_m3
```

### Model Comparison

We compare the multivariate models of the Decreasing trajectory using leave-one-out cross-validation (LOO-CV) to evaluate whether additional moderator domains improve predictive accuracy, ranking models by ELPD (with higher values indicating better performance), interpreting ΔELPD and its SE to gauge relative fit, and using optional model weights to summarize each model’s likelihood of generalizing well to new data.

```{r BRMS-model-comparison-Decreasing, results='asis', message=FALSE}
# Compare models using LOO-CV
gt_results_Decreasing <- loo_gt_table(
  model_list = list(
    models_Decreasing$model_1_bg,
    models_Decreasing$model_2_bg_trauma,
    models_Decreasing$model_3_full
  ),
  model_names = c( "M1 (Background)", "M2 (Background + Trauma)", "M3 (Background + Trauma + Methodological)"),
  response = "Decreasingn"
)

gt_results_Decreasing
```

## ⚠️ Increasing Symptom Trajectory

### Data Preparation and Correlation of Continuous Variables

Before fitting the multivariate models, data are prepared for Bayesian estimation through standardising continuous moderators and recoding categorical variables, ensuring stable estimation and interpretable coefficients.
A correlation matrix for the Increasing trajectory is used to detect highly collinear variables, with those exceeding a .90 correlation being excluded from analyses.

```{r BRMS-correlations-Increasing, results='asis', message=FALSE}
# Prepare the data (cleaning, standardizing, recoding)
df_Increasing_clean <- prepare_data_brms(df_Increasing, exclude = c("TP1_TPX", "Military"))

# Determine Collinearity
correlations_Increasing <- generate_correlation_gt(df = df_Increasing_clean, trajectory = "Increasing", outcome = "Increasing_n")
correlations_Increasing

exclude_vars_Increasing <- correlations_Increasing$exclude_vars
```

### Model Fits

The analysis of the Increasing Trajectory multivariate models distils posterior distributions by reporting each predictor’s median effect size and 95% credible interval, highlighting predictors whose intervals do not include zero. Displaying the three models together shows how effects change as additional predictors are introduced. Model diagnostics, including R-hat and effective sample sizes, confirm good convergence and lend confidence to the reliability of the posterior estimates and resulting inferences.

```{r BRMS-fitting-Increasing, results='asis', message=FALSE}
# Fit Models for Increasing Trajectory
models_Increasing <- fit_brms_trajectory_models(
  df               = df_Increasing_clean,
  success_col_name = "Increasing_n",
  cont_vars        = cont_vars,
  cat_vars         = cat_vars,
  mod_specs        = MOD_SPECS,
  background_vars  = background_vars,
  trauma_vars      = trauma_vars,
  method_vars      = method_vars,
  brms_args        = brms_args,
  exclude_vars     = exclude_vars_Increasing
)

# Produce predictor summary table
table_Increasing <- generate_predictor_table_gt(
  fitted_models = models_Increasing,
  trajectory_name = "Increasing"
)
table_Increasing

# Diagnostics for each submodel
diag_Increasing_m1 <- diagnostic_gt_brms(model = models_Increasing$model_1_bg,
                                  model_name = "M1: Background",
                                  trajectory_name = "Increasing")

diag_Increasing_m2 <- diagnostic_gt_brms(model = models_Increasing$model_2_bg_trauma,
                                  model_name = "M2: Background + Trauma",
                                  trajectory_name = "Increasing")

diag_Increasing_m3 <- diagnostic_gt_brms(model = models_Increasing$model_3_full,
                                  model_name = "M3: Background + Trauma + Methodological",
                                  trajectory_name = "Increasing")

diag_Increasing_m1
diag_Increasing_m2
diag_Increasing_m3
```

### Model Comparison

We compare the multivariate models of the Increasing trajectory using leave-one-out cross-validation (LOO-CV) to evaluate whether additional moderator domains improve predictive accuracy, ranking models by ELPD (with higher values indicating better performance), interpreting ΔELPD and its SE to gauge relative fit, and using optional model weights to summarize each model’s likelihood of generalizing well to new data.

```{r BRMS-model-comparison-Increasing, results='asis', message=FALSE}
# Compare models using LOO-CV
gt_results_Increasing <- loo_gt_table(
  model_list = list(
    models_Increasing$model_1_bg,
    models_Increasing$model_2_bg_trauma,
    models_Increasing$model_3_full
  ),
  model_names = c( "M1 (Background)", "M2 (Background + Trauma)", "M3 (Background + Trauma + Methodological)"),
  response = "Increasingn"
)

gt_results_Increasing
```

## 🩸 High Symptom Trajectory

### Data Preparation and Correlation of Continuous Variables

Before fitting the multivariate models, data are prepared for Bayesian estimation through standardising continuous moderators and recoding categorical variables, ensuring stable estimation and interpretable coefficients.
A correlation matrix for the High trajectory is used to detect highly collinear variables, with those exceeding a .90 correlation being excluded from analyses.

```{r BRMS-correlations-High, results='asis', message=FALSE}
# Prepare the data (cleaning, standardizing, recoding)
df_High_clean <- prepare_data_brms(df_High, exclude = c("TP1_TPX", "Military"))

# Determine Collinearity
correlations_High <- generate_correlation_gt(df = df_High_clean, trajectory = "High", outcome = "High_n")
correlations_High

exclude_vars_High <- correlations_High$exclude_vars
```

### Model Fits

The analysis of the High Trajectory multivariate models distils posterior distributions by reporting each predictor’s median effect size and 95% credible interval, highlighting predictors whose intervals do not include zero. Displaying the three models together shows how effects change as additional predictors are introduced. Model diagnostics, including R-hat and effective sample sizes, confirm good convergence and lend confidence to the reliability of the posterior estimates and resulting inferences.

```{r BRMS-fitting-High, results='asis', message=FALSE}
# Fit Models for High Trajectory
models_High <- fit_brms_trajectory_models(
  df               = df_High_clean,
  success_col_name = "High_n",
  cont_vars        = cont_vars,
  cat_vars         = cat_vars,
  mod_specs        = MOD_SPECS,
  background_vars  = background_vars,
  trauma_vars      = trauma_vars,
  method_vars      = method_vars,
  brms_args        = brms_args,
  exclude_vars     = exclude_vars_High
)

# Produce predictor summary table
table_High <- generate_predictor_table_gt(
  fitted_models = models_High,
  trajectory_name = "High"
)
table_High

# Diagnostics for each submodel
diag_High_m1 <- diagnostic_gt_brms(model = models_High$model_1_bg,
                                  model_name = "M1: Background",
                                  trajectory_name = "High")

diag_High_m2 <- diagnostic_gt_brms(model = models_High$model_2_bg_trauma,
                                  model_name = "M2: Background + Trauma",
                                  trajectory_name = "High")

diag_High_m3 <- diagnostic_gt_brms(model = models_High$model_3_full,
                                  model_name = "M3: Background + Trauma + Methodological",
                                  trajectory_name = "High")

diag_High_m1
diag_High_m2
diag_High_m3
```

### Model Comparison

We compare the multivariate models of the High trajectory using leave-one-out cross-validation (LOO-CV) to evaluate whether additional moderator domains improve predictive accuracy, ranking models by ELPD (with higher values indicating better performance), interpreting ΔELPD and its SE to gauge relative fit, and using optional model weights to summarize each model’s likelihood of generalizing well to new data.

```{r BRMS-model-comparison-High, results='asis', message=FALSE}
# Compare models using LOO-CV
gt_results_High <- loo_gt_table(
  model_list = list(
    models_High$model_1_bg,
    models_High$model_2_bg_trauma,
    models_High$model_3_full
  ),
  model_names = c( "M1 (Background)", "M2 (Background + Trauma)", "M3 (Background + Trauma + Methodological)"),
  response = "Highn"
)

gt_results_High
```

## 🌗 Moderate Symptom Trajectory

### Data Preparation and Correlation of Continuous Variables

Before fitting the multivariate models, data are prepared for Bayesian estimation through standardising continuous moderators and recoding categorical variables, ensuring stable estimation and interpretable coefficients.
A correlation matrix for the Moderate trajectory is used to detect highly collinear variables, with those exceeding a .90 correlation being excluded from analyses.

```{r BRMS-correlations-Moderate, results='asis', message=FALSE}
# Prepare the data (cleaning, standardizing, recoding)
df_Moderate_clean <- prepare_data_brms(df_Moderate, exclude = c("TP1_TPX", "Military"))

# Determine Collinearity
correlations_Moderate <- generate_correlation_gt(df = df_Moderate_clean, trajectory = "Moderate", outcome = "Moderate_n")
correlations_Moderate

exclude_vars_Moderate <- correlations_Moderate$exclude_vars
```

### Model Fits

The analysis of the Moderate Trajectory multivariate models distils posterior distributions by reporting each predictor’s median effect size and 95% credible interval, highlighting predictors whose intervals do not include zero. Displaying the three models together shows how effects change as additional predictors are introduced. Model diagnostics, including R-hat and effective sample sizes, confirm good convergence and lend confidence to the reliability of the posterior estimates and resulting inferences.

```{r BRMS-fitting-Moderate, results='asis', message=FALSE}
# Fit Models for Moderate Trajectory
models_Moderate <- fit_brms_trajectory_models(
  df               = df_Moderate_clean,
  success_col_name = "Moderate_n",
  cont_vars        = cont_vars,
  cat_vars         = cat_vars,
  mod_specs        = MOD_SPECS,
  background_vars  = background_vars,
  trauma_vars      = trauma_vars,
  method_vars      = method_vars,
  brms_args        = brms_args,
  exclude_vars     = exclude_vars_Moderate
)

# Produce predictor summary table
table_Moderate <- generate_predictor_table_gt(
  fitted_models = models_Moderate,
  trajectory_name = "Moderate"
)
table_Moderate

# Diagnostics for each submodel
diag_Moderate_m1 <- diagnostic_gt_brms(model = models_Moderate$model_1_bg,
                                  model_name = "M1: Background",
                                  trajectory_name = "Moderate")

diag_Moderate_m2 <- diagnostic_gt_brms(model = models_Moderate$model_2_bg_trauma,
                                  model_name = "M2: Background + Trauma",
                                  trajectory_name = "Moderate")

diag_Moderate_m3 <- diagnostic_gt_brms(model = models_Moderate$model_3_full,
                                  model_name = "M3: Background + Trauma + Methodological",
                                  trajectory_name = "Moderate")

diag_Moderate_m1
diag_Moderate_m2
diag_Moderate_m3
```

### Model Comparison

We compare the multivariate models of the Moderate trajectory using leave-one-out cross-validation (LOO-CV) to evaluate whether additional moderator domains improve predictive accuracy, ranking models by ELPD (with higher values indicating better performance), interpreting ΔELPD and its SE to gauge relative fit, and using optional model weights to summarize each model’s likelihood of generalizing well to new data.

```{r BRMS-model-comparison-Moderate, results='asis', message=FALSE}
# Compare models using LOO-CV
gt_results_Moderate <- loo_gt_table(
  model_list = list(
    models_Moderate$model_1_bg,
    models_Moderate$model_2_bg_trauma,
    models_Moderate$model_3_full
  ),
  model_names = c( "M1 (Background)", "M2 (Background + Trauma)", "M3 (Background + Trauma + Methodological)"),
  response = "Moderaten"
)

gt_results_Moderate
```
---
